---
title: "Practical Machine Learning project"
author: "Petr Lezak"
date: "2018-11-18"
output: html_document
---

The goal of this project was to create a classification model for detecting how was unilaterall dumbell biceps curl done. Classification should be done by data from accelerometers. The classification is stored in varaible classe with 5 different possible values:

- 'A' - according to the specification
- 'B' - throwing the elbows to the front
- 'C' - lifting the dumbell only halfway
- 'D' - lowering the dumbell only halfway
- 'E' - throwing the hips to the front

My goal was to achieve at least 80% accuracy necessary to pass the test. 
I have chosen multiple models with logistic regression. There are 5 models, each predicting if the classification is equal to one of 5 possible ones. Model with highest probability of positive result is chosen as prediction. I have trained models with all non-empty features from the test set and obtained accuracy cca 70% which is not enough. I have then added quadratic terms of these features (just squared, not mixed terms) and obtained accuracy about 90% which is more than enough.

Let us begin with splitting training set to trainingData and testData. Also load verification data - this is the data for the quiz.

```{r}
library(caret);

set.seed(88888);
data <- read.csv('pml-training.csv');
verification <- read.csv('pml-testing.csv');

inTrain <- createDataPartition(y = data$classe, p = 0.7, list = FALSE);
trainingData <- data[inTrain, ];
testData <- data[-inTrain, ];
```

Let us define a function train that will train a model for given classification.

```{r}
train <- function(data, classification) {
  glm(
    I(classe == classification)~
      num_window+
      roll_belt+
      pitch_belt+
      yaw_belt+
      total_accel_belt+
      gyros_belt_x+
      gyros_belt_y+
      gyros_belt_z+
      accel_belt_x+
      accel_belt_y+
      accel_belt_z+
      magnet_belt_x+
      magnet_belt_y+
      magnet_belt_z+
      roll_arm+
      pitch_arm+
      yaw_arm+
      total_accel_arm+
      gyros_arm_x+
      gyros_arm_y+
      gyros_arm_z+
      accel_arm_x+
      accel_arm_y+
      accel_arm_z+
      magnet_arm_x+
      magnet_arm_y+
      magnet_arm_z+
      roll_dumbbell+
      pitch_dumbbell+
      total_accel_dumbbell+
      yaw_dumbbell+
      gyros_dumbbell_x+
      gyros_dumbbell_y+
      gyros_dumbbell_z+
      accel_dumbbell_x+
      accel_dumbbell_y+
      accel_dumbbell_z+
      magnet_dumbbell_x+
      magnet_dumbbell_y+
      magnet_dumbbell_z+
      roll_forearm+
      pitch_forearm+
      yaw_forearm+
      gyros_forearm_x+
      gyros_forearm_y+
      gyros_forearm_z+
      accel_forearm_x+
      accel_forearm_y+
      accel_forearm_z+
      magnet_forearm_x+
      magnet_forearm_y+
      magnet_forearm_z+
      
      I(num_window^2)+
      I(roll_belt^2)+
      I(pitch_belt^2)+
      I(yaw_belt^2)+
      I(total_accel_belt^2)+
      I(gyros_belt_x^2)+
      I(gyros_belt_y^2)+
      I(gyros_belt_z^2)+
      I(accel_belt_x^2)+
      I(accel_belt_y^2)+
      I(accel_belt_z^2)+
      I(magnet_belt_x^2)+
      I(magnet_belt_y^2)+
      I(magnet_belt_z^2)+
      I(roll_arm^2)+
      I(pitch_arm^2)+
      I(yaw_arm^2)+
      I(total_accel_arm^2)+
      I(gyros_arm_x^2)+
      I(gyros_arm_y^2)+
      I(gyros_arm_z^2)+
      I(accel_arm_x^2)+
      I(accel_arm_y^2)+
      I(accel_arm_z^2)+
      I(magnet_arm_x^2)+
      I(magnet_arm_y^2)+
      I(magnet_arm_z^2)+
      I(roll_dumbbell^2)+
      I(pitch_dumbbell^2)+
      I(total_accel_dumbbell^2)+
      I(yaw_dumbbell^2)+
      I(gyros_dumbbell_x^2)+
      I(gyros_dumbbell_y^2)+
      I(gyros_dumbbell_z^2)+
      I(accel_dumbbell_x^2)+
      I(accel_dumbbell_y^2)+
      I(accel_dumbbell_z^2)+
      I(magnet_dumbbell_x^2)+
      I(magnet_dumbbell_y^2)+
      I(magnet_dumbbell_z^2)+
      I(roll_forearm^2)+
      I(pitch_forearm^2)+
      I(yaw_forearm^2)+
      I(gyros_forearm_x^2)+
      I(gyros_forearm_y^2)+
      I(gyros_forearm_z^2)+
      I(accel_forearm_x^2)+
      I(accel_forearm_y^2)+
      I(accel_forearm_z^2)+
      I(magnet_forearm_x^2)+
      I(magnet_forearm_y^2)+
      I(magnet_forearm_z^2)
    ,
    data = data, family = binomial(link = "logit")
  );
}
```

Now we uses the function train for building five different models, each for one classification.

```{r results='hide', warning=FALSE}
mA <- train(trainingData, 'A');
mB <- train(trainingData, 'B');
mC <- train(trainingData, 'C');
mD <- train(trainingData, 'D');
mE <- train(trainingData, 'E');
```

Let us define a function pred that returns vector of predictions for given data. This function applies 5 previously trained models to the data and for each sample returns classification with highest probability.

```{r}
pred <- function(data)
{
  a <- predict(mA, newdata = data);
  b <- predict(mB, newdata = data);
  c <- predict(mC, newdata = data);
  d <- predict(mD, newdata = data);
  e <- predict(mE, newdata = data);
  
  res <- vector(length = nrow(data))
  
  for (i in 1:nrow(data)) {
    m <- max(a[i], b[i], c[i], d[i], e[i]);
    
    if (a[i] == m) {
      res[i] <- 'A';
    }
    
    if (b[i] == m) {
      res[i] <- 'B';
    }
    
    if (c[i] == m) {
      res[i] <- 'C';
    }
    
    if (d[i] == m) {
      res[i] <- 'D';
    }
    
    if (e[i] == m) {
      res[i] <- 'E';
    }
  }
  
  res;
}
```

Now we run the prediction on traing data and test data and print accuracies. Also we run the classification on verification data to solve the quiz - all of them are correct.

```{r}
predictTrain <- pred(trainingData)
sum(predictTrain == trainingData$classe) / length(trainingData$classe)

predictTest <- pred(testData)
sum(predictTest == testData$classe) / length(testData$classe)

predictVerification <- pred(verification)
predictVerification
```

Finally we calculate confusion matrix for test set. Rows would be predicted classes, columns actual classes.

```{r}
classes <- c('A', 'B', 'C', 'D', 'E');
confusion <- outer(classes, classes, Vectorize (FUN = function(x, y) {(fun = sum(predictTest == x & testData$classe == y))}));

rownames(confusion) <- classes;
colnames(confusion) <- classes;

confusion
```
